---
title: "Cleaning script for palmerpenguin Data with Exploratory Plots"
author: "Allison Fisher"
date: "2023-03-02"
output: html_document
---


# Processing script (data cleaning) with code pulled in from script

This is based on the processingfile_v2.qmd template to pull in chunks of code from the R script.

# Setup

Pulling in the R script.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('processingcode.R')
```

Load needed packages.


```{r, packages, message = FALSE, warning = FALSE}
```

# Data loading
Let's load the data then make and explore the dictionary.

```{r, loaddata}
```

# Explore Data

### Take a look at the data

```{r, exploredata}
```


# Cleaning

###  Species names

Let's fix the typos in the species names to make them all consistent.

```{r, cleandata1}
```

Shorten Species just keeping the three common names "Adelie", "Gentoo", and "Chinstrap" and delete the rest of the Species character string.

```{r, shortennames}
```

### Continuous data

Replace the missing entry for `Culmen Length (mm)` to NA and coerce the variable to numeric.

```{r, cleandata2}
```

Now let's fix the Culmen Lengths which are a decimal place off

```{r, cleandata3.1}
```
### Now let's look at body mass.

Get rid of the masses too small to be adults by replacing with NA and dropping them.

```{r, cleandata4}
```

### Factors

We also want to have Species, Sex, and Island coded as a categorical/factor variable:

```{r, cleandata5}
```

# Bivariate Plots

Make some plots to check the data for outliers. We can also look at the distributions of categorical variables using ggplot!

```{r, exploratoryplots}
```

# Finalize your cleaned dataset.

I decided to drop variables studyName and Stage (they were all the same).

```{r, finalizedata}
```
# Save data

Now we make the cleaned data easy to save and save it as both .rds and .csv format

```{r, savedata}
```
# View data once more

Let's have a final look at the cleaned table, FYI

```{r, finallook}
```
---
title: "project1"
---
